---
title: "miRcore R camp TCGA Group Task"
author: "[Group Name Here]"
date: "July 11, 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

```

# miRcore summer camp 2023
© miRcore


## Disease Information

What is the significance ((background and problems) of the disease you are analyzing?



## Reading in Data

Read in the cancer miRNA expression text file that you downloaded from the Google Drive

```{r}

kirc = read.table("KIRC.txt", header = T, stringsAsFactors = F, row.names=1)
```


A. What are the dimensions of your text file?
886 100
```{r}

dim(kirc)
head(kirc)


```

B. How many Tumor and Control samples are in your dataset?

```{r}
length(grep("Control", colnames(kirc)))
length(grep("Tumor", colnames(kirc)))

```


## Plotting

A1. Make box plots for first 10 Tumor samples and first 10 Control samples using all data.

```{r}
boxplot(kirc[, c(1:10, 50:60)], main = "Boxplot of first 10 tumor and first 10 control")
```

A2. Make box plots for first 10 Tumor samples and first 10 Control samples using log2(data+1).

```{r}
log.kirc = log2(kirc+1)
boxplot(log.kirc[, c(1:10, 50:60)], main = "Log2 boxplot of first 10 tumor and first 10 control", xlab = "Sample", ylab = "Expression")
```


B. Make a scatterplot for hsa-miR-200a-3p and hsa-miR-429 using all
samples in your data file. Then compute the Pearson correlation.

```{r}


```

Check that no rows have a sum of 0 
Use rowSums() function 
eg: rowSums(dataframe[column numbers vector], na.rm=TRUE)
Remove needed rows 
Use a for loop to calculate separately the mean of each row
Calculate log2 fc
log(mean(T)/mean(C),2)
Make sure corresponding tumor and control columns are selected
Save the results!
Think about what data structure would be best
```{r}
nKIRC <- data.frame()
control_cols <- grep("Control", colnames(kirc))
tumor_cols <- grep("Tumor", colnames(kirc))

for (i in 1:nrow(kirc)) {
  if (sum(kirc[i, control_cols]) != 0 && sum(kirc[i, tumor_cols]) != 0) {
    nKIRC <- rbind(nKIRC, kirc[i, ])
  }
}

nKIRC
```

## T Tests


A. For each miRNA: calculate mean, standard deviation, and the 2-tail student t-test p-value of Tumor and Control groups. Save all results in a table.

```{r}
kircmatrix = as.matrix(kirc)
kircttest = data.frame()
for(i in 1:nrow(kirc)) {
  kircTumor = kircmatrix[i, tumor_cols]
  kircControl = kircmatrix[i, control_cols]
  kircTumorMean = mean(kircTumor)
  kircTumorSD = sd(kircTumor)
  kircControlMean = mean(kircControl)
  kircControlSD = sd(kircControl)
  pvalue = t.test(kircTumor, kircControl)$p.value
  kircttest[i, "Tumor Mean"] = kircTumorMean
  kircttest[i, "Control Mean"] = kircControlMean
  kircttest[i, "Tumor SD"] = kircTumorSD
  kircttest[i, "Control SD"] = kircControlSD
  kircttest[i, "T-Test"] = t.test(kircTumor, kircControl)
  kircttest[i, "pvalue"] = pvalue
}
rownames(kircttest) = rownames(kircmatrix)
kircttest

```



B.  Find 50 most significantly different miRNA names between tumor and control samples and save their names with means, standard deviations for tumor and control samples and p-values.

```{r}
order(kircttest$pvalue)
statTableOrder <- kircttest[order(kircttest$pvalue), ]
sig_50 <- statTableOrder[1:50, ]
head(sig_50)


```


C. Make a heatmap of 50 most signficantly differently expressed miRNAs (row clustering only). Use the colors darkblue,
white and red2. [USA!USA!USA!].

```{r}
logKIRC = log2(kircmatrix+1)
sigKIRC = logKIRC[order(kircttest$pvalue)[1:50] , ]

library(pheatmap)

pheatmap(sigKIRC, scale = "row", cluster_cols = F, border_color = NA, main = "Sig Expressions", cellwidth = 3, cellheight = 3,
color = colorRampPalette(c("darkblue", "white", "red2"))(256))
```

## Correlations

A. Find correlations between the most significant miRNA with 29 other miRNAs for 1) all samples.
```{r}

```


B. Among A. can you find some patterns between miRNAs? What are your findings? (example: to find the most correlated miRNA with the most significant miRNA)


## Random Forest (Bonus)

A. Create a random forest using the steps from the random forest .html file to classify your data. What is your accuracy?


B. Change variables in the random forest run and check if the results are different. What are the parameters that you can achieve a better accuracy?


## Analysis

### miRNAs


What are your most significant miRNAs? Are they up-regulated or down-regulated? Are they highly correlated with each other?

```{r}

```


### Genes

A. What are the common gene targets of your significant miRNAs?


B. Are any of your genes targeted by these top miRNAs known mutations of the disease you are researching?


## Conclusion

What do your findings suggest to you about potential diagnosis/treatment targets for the disease you are researching?



## Optional


A. Make a heatmap of all data (row clustering only). Use the colors darkblue, white and red2. [USA!USA!USA!].


B. find all correlations among the 30 most significant miRNAs for
1) all samples, 2) all tumor samples, 3) all control samples.
