---
title: "2R"
author: "Prasanna Padmanabham"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The learning objectives for this lecture is to:
1) learn how to read files correctly
2) Review Indexing, Which() and grep()
3) Create new columns, and column operations

### Reading Files
Our goal is to load the small_data.txt file into Rstudio so we can analyze the data and create really cool visuals

Small data has 20 rows and 20 columns

The function we will be using is the read.table()
```{r}
smalldata= read.table("small_data.txt")
```

Let's check to see if our data was loaded correctly: 
```{r}
dim(smalldata)
head(smalldata)
str(smalldata)
```

Did it work? 


Let's try again: 
```{r}
smalldata = read.table("small_data.txt", header=T, stringsAsFactors = F)
dim(smalldata)
head(smalldata)
str(smalldata)
```

Did it work? 

3rd time's the charm: 
```{r}
smalldata = read.table("small_data.txt", header=T, stringsAsFactors = F, row.names = 1)
dim(smalldata)
head(smalldata)
str(smalldata)
colnames(smalldata)
row.names(smalldata)
```

#### Breakout Session
1) Make sure you read small_data.txt correctly
2) Update your R markdown files and your cheat sheet
3) What do you think the functions colnames(smalldata) and row.names(smalldata) do? 
4) Make a group Zoom background


### Indexing and Subseting

Let's review how to index dataframes (and vectors) using our first data table, smalldata!

How can we find the value associated with the 20 row and 15 column? 
```{r}
smalldata[20,15]
smalldata[1:5, 1:10]#Let's find the first 5 rows and the first 10 columns
smalldata[1:5,]#returns the first 5 rows and all columns
smalldata["mir_a",]#returns all values in the "miR_a" row
smalldata[,"Tumor_1"]#returns all values in Tumor_1 column

#If you want to subset multiple rows you can even do
smalldata[c("miR_a, miR_e"), c(1,5,20.15)]
#or even
```

But what if we don't know what specific row we are interested in  or the specific index numbers associated with ALL tumors and ALL controls? 

### Grep
type ?grep() into your console, what do you think it does?
grep(pattern, x)

pattern: "character string...to be matched in the given character vector"
x: "a character vector where matches are sought"

Let's find the column id's associated with all controls in smalldata
```{r}
colnames(smalldata)
grep("Control", colnames(smalldata))
smalldata[,grep("Control", colnames(smalldata))]
```
Let's select row names with specific words such as "miR_b"
```{r}
row.names(smalldata)
grep("miR_b", row.names(smalldata))
smalldata[grep("miR_b", row.names(smalldata)),]
```

Did this actually work? 
```{r}
head(smalldata)
```

Yes! miR_b is the second row and all the control data is located in columns 11-20!

What if we want to subset rows or columns that contain specific values?
Let's find all the miR_a values over 20,000
```{r}
which(smalldata["miR_a",] >= 20000)
head(smalldata)
```

All in all, grep and which can be used to find specific patterns in dataframes or vectors
(switch to powerpoint here)

#### Breakout Room task: 
Update your Rmd/cheatsheet on grep

Get column ids for “Tumor” samples
Select all values for Tumor samples from smalldata
```{r}
colnames(smalldata)
grep("Tumor", colnames(smalldata))
smalldata[,grep("Tumor", colnames(smalldata))]

```


Perfect your group zoom background

### Creating a new table from small_data.txt
We want a new table with just Tumor_1 and Control_1 values
```{r}
Tumor1 = smalldata[, "Tumor_1"]
Control1 = smalldata[, "Control_1"]
Tum1_Con1 = cbind.data.frame(Tumor1, Control1)
```

Does our new dataframe look the way we want it to look?
```{r}
row.names(Tum1_Con1) = row.names(smalldata)
head(Tum1_Con1)
```
```{r}

```



Now let's make a new col with the difference in expression between all the miRs for Tumor1 and Control1
```{r}
Tum1_Con1$Difference = Tum1_Con1[,1]- Tum1_Con1[,2]
```

Let's save our table now!
```{r}
write.table(Tum1_Con1, file = "Sample1_Data.txt", sep = "\t", col.names=NA)
dir()
```


#### Breakout Room Tasks: 
1) Catch up, you should have the correct Tum1_Con1 and smalldata frames
2) Write commands to achieve the following: 
  a) Assign new row.names to Tum1_Con1 using the miRNA names from smalldata
```{r}
row.names(Tum1_Con1) = row.names(smalldata)
```
  
  
  b) Calculate value of Tumor1/Control1 for each miRNA and assign all calculated values to a
  new column called Ratio in the Tum1_Con1 dataframe
```{r}
Tum1_Con1$Ratio = Tum1_Con1[,1]/Tum1_Con1[,2]
head(Tum1_Con1)
```
  
  
  c) Find the names of miRNA's whose Tumor1/Control1 values are less than 1 (aka down
  regulated miRNAs in Tumor1)
```{r}
row.names(Tum1_Con1[which(Tum1_Con1[,"Ratio"] < 1),])
head(Tum1_Con1)
```
  

