---
title: "Day1"
author: "Inhan Lee"
date: "7/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## copyright (c)miRcore
## miRcore R camp day 1

### basic calculation

```{r}

```

## vector

Take notes here!

```{r}
vec1 = 1:5
vec2 = c(2, 4, 6, 8, 10)
vec3 = seq(from=0, to=20, by=4)
```


### scan function
function scan() gets input from keyboards through Console window and form a vector. Carriage return separates the element. Two carriage return close the scan function. Default data type: character.

change input data type to numeric: scan(what=numeric()) 
1) run the code: vec4 <- scan(what=numeric())
2) go to Console section (lower left section) and activate cursor there
3) type numbers 20 30 15 22 18 from Console with carriage returns
4) end: one more carriage return
```{r}
vec4 = scan(what=numeric())
```


```{r}
vec2
vec2[1]
vec2[1] + vec2[2] + vec2[3] + vec2[4] + vec2[5]
sum(vec2)
```



### element of vector []
```{r}
sum(vec2) #find the sum of the vector
mean(vec2) #find the mean of the vector
sd(vec2) #find the standard deviation of the vector


```
Check 
```{r}

```

### vector has the same data type
```{r}
mix = c(1, 2, 3, 4, "five")
mix
str(mix)
mix2 = as.numeric(mix)
str(mix2)
mean(mi2, na.rm=TRUE)
```

### special character
NA, NaN, 
```{r}
mean(mix)
mean(mix2)
```

## which
```{r}
vec2 == 10
?which
which(vec2 == 20)
```

## matrix


Method 1 of creating a matrix by using matrix() function
```{r}
M1 = matrix(c(3, 7, 4, 2, 5, 9), nrow=2, ncol=3)
M1
# prints the matrix
```

Method 2 of creating a Matrix (by combining vectors)
```{r}
a.vec = c(2, 4, 6)
b.vec = c(3, 5, 7)
M2 = cbind(a.vec, b.vec)
M2
M3 = rbind(a.vec, b.vec)
```

### matrix element
M1[rownumber, colnumber]
```{r}
M1 = matrix(c(3, 7, 4, 2, 5, 9), nrow=2, ncol=3)
M1[1, 1]
M1[2, 1]
M1[1,]
M1[,2]
M1[2, 3]
```
MAT12[rownumber, colnumber]
```{r}
MAT12 = cbind(vec2, vec3)
MAT12[3, 2]
which(MAT12 ==8, arr.ind =T)
```


## data frame

Let's make a data frame with:
column1: numbers between 1 through 14
column2: letters a - d
coloumn3: group 1001
```{r}
numbers = 11:14 #vector 1
letters = c("a", "b", "c", "d") #vector 2
group = as.factor(c(1,0,0,1)) #vector 3

mydata = cbind.data.frame(numbers, letters, group) #cbind three vectors
str(mydata) #what is the structure of mydata?
rownames(mydata) = c("A", "B", "C", "D") #let's add rownames
mydata #what does mydata look like now?
colnames(mydata) #what are the column names of mydata
rownames(mydata) #what are the rownames of mydata
```
```{r}
mydata[,2]
mydata[, "letters"]
mydata$letters
```


Pattern match
```{r}
head(mydata)
tail(mydata)

mydata[1:3]
mydata[c(3, 2, 1),]

mydata[,1:3]
mydata[,c(3, 2, 1)]
```

```{r}
mydata[mydata[,2] == "b",]
```

```{r}
mydata[2, 3]
mydata[1,]
mydata[1:3, 2]
mydata[1:3, "numbers"]
mydata[mydata[2] == "13"]
```

