---
title: "ggplot_basic"
author: "Inhan Lee"
date: "6/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Open your Rstudio and load the blank file.

Today we will use ggplot to make high-quality figures. 
The graphic philosophy: We define a playing ground (data frame), with an x-axis and a y-axis and then we can add layers over it. 

We do asesthetic mapping of variables in the data to visual properties.

## Install package and setup

```{r}
install.packages("ggplot2")
library(ggplot2)
```


## Read data

ggplot needs dataframe to read x-values and y-values.

#### We will also prepare a log transformed dataframe of smalldata


```{r}
smalldata = read.table("small_data.txt", header = T, row.names = 1)
smalldata_log = log2(smalldata+1)
 
```

## ggplot first define the data range (x and y)
#### x and y: both columns of a dataframe

```{r}
ggplot(smalldata_log, aes(x = Tumor_1, y= Control_1)) +
  geom_point(aes(color = Tumor_1))
ggplot(smalldata_log,aes(x=Tumor_1, y=Control_1)) + geom_point(aes(color=Tumor_1, size=Tumor_1)) + scale_color_gradient(low = "white", high = "black")
smalldata_log$cutoff=smalldata_log$Tumor_1>10
ggplot(smalldata_log,aes(x=Tumor_1,y=Control_1))+
  geom_point(aes(color=cutoff))
ggplot(smalldata_log,aes(x=Tumor_1,y=Control_1))+
  geom_point()+
  geom_point(aes(x=Tumor_2,y=Control_2),color="red")+
  xlab("Tumor expression")+
  ylab("Control expression")
```

## Second add the graph type
geom_point() is used to create scatterplots

```{r}
values=c(rnorm(20,mean=10,sd=2),rnorm(20,mean=12,sd=4),rnorm(20,mean=5,sd=1))

group=c(rep("A",20),rep("B",20),rep("C",20))

gg_example=cbind.data.frame(values,group)

ggplot(data = gg_example,aes(x=group, y=values)) +
  geom_boxplot((aes(color = group)))

ggplot(data = gg_example,aes(x=group, y=values)) +
  geom_point() +
  geom_boxplot((aes(fill = group)))

ggplot(data=gg_example,aes(x=group,y=values))+
  geom_boxplot(aes(fill=group))+
  geom_point()
  
```


## We can change color and size of the points
Let's use red for points with size 3.

```{r}



```

## Tasks

1. Use log transformed data range (x and y) to define the base layer of x-values and y-values

```{r}


```

2. Use scatter plot to define another layer of plot type

```{r}


```


3. Change the color and size of the points

```{r}

```


4. additional layer challenge: add abline to current graph
```{r}


```


5. Change the color and thickness of the abline


```{r}

```





