---
title: "Iteration and Conditional"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Iterations
Prompt: Let's say I want to type the statement "My favorite number is 1" with 100 different numbers. How do I do this? 

First let's try printing the statement manually: 
```{r}

```

I don't know about you, but I'm starting to get a bit bored of copying, pasting and changing the number...let's see if we can utalize the for loop

```{r}
for (i in 1: 10) {
  print(paste("My favorite number is", i, sep = " "))
}
```

The first part dictates how many times the loop run for. The variable i starts with a value of 1 and will increase it's value after it finishes the loop.

i is like a temporary variable; it disappears after the for loop is done running

Let's apply this to our printing problem: 
```{r}

```

(SHOW SLIDE 8)

### Conditional
What if I only like even numbers? How can I reprogram the earlier program so only even numbers print? 
```{r}
for(i in 1:20)
{
  if(i%%2 == 0)
  {
      print(paste("my favorite number is", i, sep=" "))
  }
  else 
  {
    print(paste("my favorite number is NOT", i, sep=" "))
  }
}
```


(I'm GOING TO BE DOING LOTS OF TYPING -- PLEASE TRY TO FOLLOW ALONG THE LOGIC!)
(DO not copy this example)
Example #2, let's make a vector that creates a vector of headers that go like GroupA_1, GroupB_1, GroupA_2, GroupB_2...GroupA_10, GroupB_10
```{r}

```


(DO not copy this example)
Two problems: We want to start with GroupA and we need to get rid of the 0.5
Let's fix the starting with GroupA instead of GroupB problem first

```{r}

```



(FINAL ANSWER)
Wait, we forgot to change the index numbers when we do the Headers[i] function lol
Also, Group B is 0.5 greater than what is supposed to be

```{r}

```

Great we did it! 

EXAMPLE 3: Let's make a vector containing the pvalues from the ttests of smalldata tumors vs. controls: 
```{r}
#Show how code works for one instance
#Add a for loop here: 
```



### Breakout Room Task
Your turn: 
Go to your Group!
Using small data
1. Change the smalldata to a matrix.
```{r}

```

2. Make an empty data.frame named ttest.
```{r}
ttest = data.frame()
```

3. For the first gene (hint: first row) 
      1. calculate t.test p-value between expression values all tumor samples (hint: first 10 samples and the 
         values need to be numeric) and all control samples (change from your earlier task)
```{r}
Tumor = smalldataM[1,1:10]
Control = smalldataM[1,11:20]
t.test(Tumor, Control)
if(t.test(Tumor, Control)$p.value < 0.0001) {
  ttest = rbind(ttest, smalldata[1,])
}
```
         
      2. If p-value is less than 0.0001, add to the expression values in the ttest (hintt: use rbind 
         function).
4. Do the step 3 for the remaining genes (total 19) using a for loop
```{r}
for(i in 1:nrow(smalldataM)) {
  pvalue = t.test(smalldataM[i, 1:10], smalldataM[i, 11:20]) $p.value
  if(pvalue < 0.001) {
    ttest = rbind(ttest, smalldata[i,])
  }
}
head(ttest)
```

5. Add a column that checks the ratio between means for the tumors and control (tumor_mean/control_mean) (to the ttest dataframe)
```{r}
ttestM = as.matrix(ttest)
for(i in 1:nrow(ttest)) {
  tumorMean = mean(ttestM[i, 1:10])
  controlMean = mean(ttestM[i, 11:20])
  ttest[i, "Ratio"] = tumorMean/controlMean
}
ttest

for(k in 1:nrow(smalldata))
{
  tumor_mean = mean(t(smalldata[k,1:10]))
  control_mean = mean(t(smalldata[k,11:20]))
  smalldata[k,"Ratio"] = tumor_mean/control_mean
}

smalldata
```

6. Create a new data frame called updata. Add the data for all the miRs where the tumor mean is HIGHER than the control mean among miR with p<0.0001.
```{r}
updata = smalldata[which(smalldata$Ratio >= 1)]
updata
```

