---
title: "3_1 ttest"
author: "Prasanna Padmanabham"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## T-test
First, we need generate data that we can analyze
```{r}
par(mfrow=c(1,2))
set.seed(1)
vec1 = rnorm(10, mean=10, sd=2)
boxplot(vec1, ylim = c(0,15))

set.seed(2)
vec2 = rnorm(10, mean=10, sd=2)
boxplot(vec2, ylim = c(0,15))
```


Let's visualize the data we just generated

Now, let's run a ttest 
But before we do, how confident are that these distributions are similar (green check) or not similar (red cross)

```{r}

```

what if we just want the p.value, standard deviation or just the mean?
```{r}
t.test(vec1, vec2)$p.value
mean(vec1)
mean(vec2)
sd(vec1)
sd(vec2)
```



Let's try this with small data! Run a t.test on the miR_a values for Tumor samples and Control samples
* if you forgot how use the functions for the mean (mean()) and standard deviation (sd()), feel free to use google or the help feature on R!

* R is a little quirky in that it can't find the mean of a row from a dataframe. To solve this, change it to a matrix (as.matrix()) or a vector (as.vector())

1. What are mean and standard deviation of miR_a values of Tumor samples?
```{r}
smalldataM = as.matrix(smalldata)
miRaTumor = smalldataM["miR_a", 1:10]
mean(miRaTumor)
sd(miRaTumor)
```

2. What are mean and standard deviation of miR_a values of Control samples?
```{r}
miRaControl = smalldataM["miR_a", 11:20]
mean(miRaControl)
sd(miRaControl)
```

3. What is the p-value of student t-test between miR_a values of Tumor samples and miR_a values of Control samples?
```{r}
t.test(miRaTumor, miRaControl)
```


4. Are the miR_a values of Tumor samples different from the miR_a values of Control samples? How confident are you?

Very confident that they are different

5. What is the log2 value of mean(miR_a of Tumor)/mean(miR_a of Control)?
```{r}
log2(mean(miRaTumor)/mean(miRaControl))
```

6. Is miR_a upregulated or downregulated in Tumor compared to Control? 

downregulated
